Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-23 paper/sec-analysis.tex
Completed revision of paper

\subsection{Impact of pile-up}
\label{sec:pileup}

Now we turn to discuss how the description of kinematic
distributions for $hh\to 4b$ signal
and background processes are
modified in the presence of pile-up.
%
To study the impact of PU,
Minimum Bias events have been generated
with {\tt Pythia8}, and then
superimposed to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios,
one with a mean number of
PU vertices per bunch crossing of $\la n_{\rm PU}\ra=80$,
and another
with $\la n_{\rm PU}\ra=150$.
%
While the latter is closer to the HL-LHC forecasts,
we adopt $\la n_{\rm PU}\ra=80$ as our baseline,
given that it
is beyond the scope of this paper to fully optimize
the PU subtraction.

>>>>

\subsection{Impact of pileup}
\label{sec:pileup}

Now we turn to discuss how the description of kinematic
distributions for $hh\to 4b$ signal
and background processes are
modified in the presence of pileup.
%
To study the impact of PU,
Minimum Bias events have been generated
with {\tt Pythia8}, and then
superimposed to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios,
one with a mean number of
PU vertices per bunch crossing of $\la n_{\rm PU}\ra=80$,
and another
with $\la n_{\rm PU}\ra=150$.
%
While the latter is closer to the HL-LHC forecasts,
we adopt $\la n_{\rm PU}\ra=80$ as our baseline,
given that it
is beyond the scope of this paper to fully optimize
the PU subtraction.

### 2015-12-23 paper/sec-mva.tex
Completed revision of paper

In this section we study how the MVA results are modified
when the analysis is performed including PU.
%
The cut-based analysis and the subsequent
MVA optimization have been performed using the same
settings as in the case without PU.
%
In Table~\ref{table:cutflowMVA_PU} we provide the results
  at the end of the cut-based analysis,
  for the case with $\la n_{\rm PU}\ra=80$ with SK+Trim
  subtraction,
  as well as the post-MVA results for the
  corresponding optimal values of $y_{\rm cut}$.
  %
  Additionally we quote in
  both cases the number of signal and
    background events expected
    at the HL-LHC with $\mathcal{L}=3$ ab$^{-1}$
    and the values of $S/\sqrt{B}$ and $S/B$

We observe that the pre-MVA 
signal significance is close
to the case without PU for the three categories.
%
We now find values for $S/\sqrt{B}$ of 0.9, 0.7 and 1.5, in the resolved,
intermediate and boosted categories, respectively, to be compared
with the corresponding values without PU, namely 1.0, 0.9 and 1.0.
%
Likewise, the number of selected
signal events in each category, at the
end of the cut-based analysis, is only mildly affected
by PU.
%
The slight pre-MVA improvement in $S/\sqrt{B}$ for the
boosted case arises from a reduction in the number
of background events that are classified in this category
as compared to the case without PU.

Once the MVA is applied, the signal significance in the 
resolved, intermediate and boosted
categories increases to 4.6, 4.1 and 3.7 respectively,
to be compared with 4.5, 4.8 and 5.4 in the case
without PU.
%
Therefore, the post-MVA effects of PU on $S/\sqrt{B}$ is
a moderate degradation of the boosted and intermediate categories,
while the resolved category is largely unchanged.
%
We also observe that, due
to the MVA, the
signal over background ratio is increased from 0.02\%, 0.08\% and
0.2\% up to 1\%, 4\% and 2\% in the resolved, intermediate
and boosted categories respectively.
%
This indicates that while this measurement is still highly challenging,
requiring a careful extraction of the QCD
background from the data, it is certainty within reach.

@@

In Fig.~\ref{fig:nev2_PU}
we show the number of signal and background events that
are expected at the HL-LHC as a function of
$y_{\rm cut}$, as well as the corresponding ROC curve.
%
The slight degradation of the boosted category in the case
of PU can be seen by comparing with the corresponding
results without PU in Fig.~\ref{fig:nev2}.
%
In Fig.~\ref{fig:sb_mva_PU} we show the signal significance,
$S/\sqrt{B}$, and the signal over background ratio,
$S/B$, accounting now for the effects of PU.
%
The corresponding results in the case without PU were shown in
Fig.~\ref{fig:sb_mva}.
%
As can be seen, the MVA-driven enhancement remains robust in the
presence of PU.
%
While
there is some degradation as compared to the case
without PU,
it is still possible to
achieve a signal significance of
 $S/\sqrt{B}$ between 3.0 and 4.5, depending on the
specific choice of $y_{\rm cut}$,
separately for the three categories.
%
We conclude that the qualitative conclusions drawn
in the case without PU also hold when the analysis
is performed in an environment including pileup.
%
Since no specific effort has been performed to
optimize PU subtraction, for instance by tuning the values
of the patch length $a$ in {\tt SoftKiller}
and of the $p_T$ threshold in jet trimming.
we believe that
there could be still room for further  improvement.

>>>>

In this section we study how the MVA results are modified
when the analysis is performed including significant PU.
%
The cut-based analysis and the subsequent
MVA optimization have been performed using the same
settings as in the case without PU.
%
In Table~\ref{table:cutflowMVA_PU} we provide the results
  at the end of the cut-based analysis,
  for the case with $\la n_{\rm PU}\ra=80$ with SK+Trim
  subtraction,
  as well as the post-MVA results for the
  corresponding optimal values of $y_{\rm cut}$.
  %
  Additionally we quote in
  both cases the number of signal and
    total background events expected
    at the HL-LHC with $\mathcal{L}=3$ ab$^{-1}$
    and the values of $S/\sqrt{B}$ and $S/B$

We observe that the pre-MVA 
signal significance is close
to the case without PU for the three categories.
%
We now find values for $S/\sqrt{B}$ of 0.9, 0.7 and 1.5, in the resolved,
intermediate and boosted categories, respectively, to be compared
with the corresponding values without PU, namely 1.0, 0.9 and 1.0.
%
The number of selected
signal events in each category, at the
end of the cut-based analysis, is only mildly affected
by PU.
%
The slight pre-MVA improvement in $S/\sqrt{B}$ for the
boosted case arises from a reduction in the number
of background events that are classified in this category
as compared to the case without PU.

Once the MVA is applied, the signal significance in the 
resolved, intermediate and boosted
categories increases to 4.6, 4.1 and 3.7 respectively,
to be compared with the corresponding values
without PU, namely 4.5, 4.8 and 5.4.
%
Therefore, the post-MVA effect of PU on $S/\sqrt{B}$ is
a moderate degradation of the boosted and intermediate categories,
while the resolved category is largely unchanged.
%
We also observe that, due
to the MVA, the
signal over background ratio is increased from 0.02\%, 0.08\% and
0.2\% up to 1\%, 4\% and 2\% in the resolved, intermediate
and boosted categories respectively.
%
This indicates that while this measurement is still highly challenging,
requiring a careful extraction of the QCD
background from the data, it is certainty within reach.

@@

In Fig.~\ref{fig:nev2_PU}
we show the number of signal and background events that
are expected at the HL-LHC as a function of
$y_{\rm cut}$, as well as the corresponding ROC curve.
%
The slight degradation of the boosted category in the case
of PU can be seen by comparing with the corresponding
results without PU in Fig.~\ref{fig:nev2}.
%
In Fig.~\ref{fig:sb_mva_PU} we show the signal significance,
$S/\sqrt{B}$, and the signal over background ratio,
$S/B$, accounting now for the effects of PU.
%
The corresponding results in the case without PU were shown in
Fig.~\ref{fig:sb_mva}.
%
As can be seen, the MVA-driven enhancement remains robust in the
presence of PU.
%
While
there is some degradation as compared to the case
without PU,
it is still possible to
achieve a signal significance of
 $S/\sqrt{B}$ between 3.0 and 4.5, depending on the
specific choice of $y_{\rm cut}$,
separately for the three categories.
%
Therefore, the qualitative conclusions drawn
in the case without PU also hold when the analysis
is performed in an environment including pileup.
%
Since no specific effort has been performed to
optimize PU subtraction, for instance by tuning the values
of the patch length $a$ in {\tt SoftKiller}
and of the $p_T$ threshold in jet trimming.
we believe that
there could be still room for further  improvement.
